---
title: "EDA Workshop"
author: "Lijiayu Deng, u8178636"
date: "Today's date"
output:
    html_document:
        toc: true
        toc_depth: 4
        theme: cosmo
        number_sections: false
        toc_float: true
        highlight: pygments
        fig_width: 8
        fig_height: 4
---

# **Task 1**: Loading the data

```{r}
library(tidyverse)
data <- read_csv("raw_data/data.csv")
data
```

### Task 1 Questions
```{r}
nrow(data)
ncol(data)
```

```{r}
data |> summarise(
  min_steps = min(steps, na.rm = TRUE),
  max_steps = max(steps, na.rm = TRUE),
  all_positive = all(steps > 0, na.rm = TRUE)
)

data |> summarise(
  min_BMI = min(BMI, na.rm = TRUE),
  max_BMI = max(BMI, na.rm = TRUE),
  all_positive = all(BMI > 0, na.rm = TRUE)
)

unique(data$sex)
```

When I check the data, the minimum steps is 0, which means there are some missing values in the data, probably because some gorillas were not tracked for a whole day. The BMI is all positive, which makes sense because BMI cannot be negative. The sex variable has two unique values, M and F, which makes sense.

# **Task 2**: Null Hypothesis Significance Testing

- We assume that the null hypothesis is that there is no difference in the average number of steps taken between male  and female gorillas.

- We can use a two-sample t-test to compare the means of the two groups

```{r}
# Calculate from the data 
data |> 
  group_by(sex) |> 
  summarise(
    n = n(),
    mean_steps = mean(steps, na.rm = TRUE),
    sd_steps = sd(steps, na.rm = TRUE)
  )
```

```{r}
t_test_result <- t.test(
  steps ~ sex, 
  data = data, 
  var.equal = TRUE
)
t_test_result
```

```{r}
# Calculate effect size (Cohen's d)
male_steps <- data |> filter(sex == "male") |> pull(steps)
female_steps <- data |> filter(sex == "female") |> pull(steps)

n1 <- length(male_steps)
n2 <- length(female_steps)
mean1 <- mean(male_steps, na.rm = TRUE)
mean2 <- mean(female_steps, na.rm = TRUE)
sd1 <- sd(male_steps, na.rm = TRUE)
sd2 <- sd(female_steps, na.rm = TRUE)

pooled_sd <- sqrt(((n1-1)*sd1^2 + (n2-1)*sd2^2) / (n1 + n2 - 2))
cohens_d <- (mean1 - mean2) / pooled_sd

cohens_d
```

### Key Results

- **Test used**: We use the two-sample t-test to compare the means
- **Effect size**: 0.48 which is a medium effect size, indicating a meaningful difference between the two groups
- **95% Confidence Interval of the difference**: [-1372.3, -928.2] steps
- **p-value**: < 2.2e-16

### Conclusions

**Scientific paper:**
A two-sample t-test revealed a statistically significant difference in daily step counts between male and female gorillas (t(1784) = -10.16, p < 0.001, Cohen's d = 0.48, 95% CI [-1372.3, -928.2]), with males walking on average 1150 more steps per day than females.

**Zoo sign:**
Male gorillas walk about 1,150 more steps per day than females — a meaningful difference supported by strong statistical evidence!





### Task 2 Questions

# **Task 3**: Exploratory Data Analysis

### Task 3 Questions